---
title: "Advanced Usage"
author: "esqLABS"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load}
library(shiny)
library(esq.handsontable)
```

## Context Menu

The table includes a built-in right-click context menu (insert row above /
below, remove row, undo / redo, copy, clear selection, clear column).

### Tracking Row Changes

Compare row counts before and after each edit. The helper below is the
logic an `observeEvent()` would call:

```{r row-change-helper}
classify_row_change <- function(old_count, new_count) {
  if (new_count > old_count) "added"
  else if (new_count < old_count) "removed"
  else "unchanged"
}

classify_row_change(5, 6)
classify_row_change(5, 4)
classify_row_change(5, 5)
```

```{r context-menu, eval = FALSE}
server <- function(input, output, session) {
  row_count <- reactiveVal(5)

  observeEvent(input$my_table_edited, {
    data <- jsonlite::fromJSON(input$my_table_edited)
    change <- classify_row_change(row_count(), nrow(data))
    if (change == "added")   showNotification("Row added")
    if (change == "removed") showNotification("Row removed")
    row_count(nrow(data))
  })
}
```

## Data Validation

Define a pure validator function that you can both run in a vignette and
call from a shiny `observeEvent()`:

```{r validation-fn}
validate_data <- function(data) {
  errors <- character()

  if (any(is.na(data$name) | data$name == "")) {
    errors <- c(errors, "Name required")
  }

  if (any(data$quantity < 0, na.rm = TRUE)) {
    errors <- c(errors, "Quantity must be >= 0")
  }

  errors
}

# A clean dataset returns no errors
clean <- data.frame(
  name = c("A", "B"),
  quantity = c(1, 2),
  stringsAsFactors = FALSE
)
validate_data(clean)

# A dirty dataset reports both rules
dirty <- data.frame(
  name = c("", "B"),
  quantity = c(-1, 2),
  stringsAsFactors = FALSE
)
validate_data(dirty)
```

```{r validation, eval = FALSE}
server <- function(input, output, session) {
  observeEvent(input$my_table_edited, {
    data <- jsonlite::fromJSON(input$my_table_edited)
    errors <- validate_data(data)

    if (length(errors) > 0) {
      showNotification(
        paste("Errors:", paste(errors, collapse = ", ")),
        type = "warning"
      )
    }
  })
}
```

## Export Data

A small dataset to demonstrate CSV / JSON export:

```{r export-data}
table_data <- data.frame(
  product = c("Widget", "Gadget"),
  price = c(29.99, 49.99),
  stringsAsFactors = FALSE
)

csv_path <- tempfile(fileext = ".csv")
write.csv(table_data, csv_path, row.names = FALSE)
file.exists(csv_path)

json_path <- tempfile(fileext = ".json")
jsonlite::write_json(table_data, json_path, pretty = TRUE)
readLines(json_path, n = 3)
```

The same logic plugged into a `downloadHandler()`:

```{r csv, eval = FALSE}
ui <- fluidPage(
  esq_tableInput("data_table", data = table_data,
                 columns = list(list(name = "product", type = "text"),
                                list(name = "price", type = "numeric"))),
  downloadButton("download", "Download CSV")
)

server <- function(input, output, session) {
  current_data <- reactiveVal(table_data)

  observeEvent(input$data_table_edited, {
    current_data(jsonlite::fromJSON(input$data_table_edited))
  })

  output$download <- downloadHandler(
    filename = function() paste0("data_", Sys.Date(), ".csv"),
    content = function(file) write.csv(current_data(), file, row.names = FALSE)
  )

  output$download_json <- downloadHandler(
    filename = function() paste0("data_", Sys.Date(), ".json"),
    content = function(file) jsonlite::write_json(current_data(), file, pretty = TRUE)
  )
}
```

## Auto-Save Pattern

```{r autosave, eval = FALSE}
server <- function(input, output, session) {
  auto_save <- debounce(reactive(input$my_table_edited), 2000)

  observe({
    data_json <- auto_save()
    req(data_json)

    data <- jsonlite::fromJSON(data_json)
    saveRDS(data, "auto_save.rds")

    showNotification("Auto-saved", duration = 1)
  })
}
```

## Integration with plotly

```{r plotly, eval = FALSE}
library(plotly)

ui <- fluidPage(
  fluidRow(
    column(6, esq_tableInput("data", data = initial_data, columns = cols)),
    column(6, plotlyOutput("chart"))
  )
)

server <- function(input, output, session) {
  table_data <- reactive({
    if (!is.null(input$data_edited)) {
      jsonlite::fromJSON(input$data_edited)
    } else {
      initial_data
    }
  })

  output$chart <- renderPlotly({
    plot_ly(table_data(), x = ~category, y = ~value, type = "bar")
  })
}
```

## Custom CSS Styling

Build the styling chunk with `htmltools` so it can be inspected before being
dropped into a shiny `tags$head()`:

```{r css}
custom_css <- htmltools::tags$style(htmltools::HTML("
  .esq-handsontable th {
    background: #f5f5f5;
    font-weight: 600;
  }
  .esq-handsontable td.htDimmed {
    background: #fafafa;
    color: #999;
  }
  .esq-handsontable td.htInvalid {
    background: #ffebee !important;
  }
"))

inherits(custom_css, "shiny.tag")
```

```{r css-app, eval = FALSE}
ui <- fluidPage(
  tags$head(custom_css),
  esq_tableInput("styled_table", data = my_data, columns = my_columns)
)
```

## Keyboard Shortcuts

Built-in Handsontable shortcuts:

| Shortcut | Action |
|----------|--------|
| Arrow keys | Navigate cells |
| Enter | Edit / Move down |
| Tab | Move right |
| Shift+Tab | Move left |
| Ctrl+C | Copy |
| Ctrl+V | Paste |
| Ctrl+Z | Undo |
| Ctrl+Y | Redo |
| Delete | Clear cell |
| Escape | Cancel editing |

## Error Handling

A safe parser that returns `NULL` on bad JSON:

```{r error-parse}
safe_parse <- function(json) {
  tryCatch(
    jsonlite::fromJSON(json),
    error = function(e) NULL
  )
}

safe_parse('[{"a": 1}]')
safe_parse("not valid json")
```

```{r errors, eval = FALSE}
observeEvent(input$my_table_edited, {
  data <- safe_parse(input$my_table_edited)
  if (is.null(data)) {
    showNotification("Error parsing data", type = "error")
    return()
  }
  # Process data...
})
```

## Troubleshooting

**Table not rendering?**

- Check package installation
- Verify data frame structure

**Dropdowns empty?**

- Ensure `source` is a character vector
- Check for typos in column names

**Updates not working?**

- Verify `inputId` matches
- Check session is passed correctly

**Performance issues?**

- Reduce row count (use pagination)
- Use simpler column types
- Minimize columns
