---
title: "Advanced Usage"
author: "esqLABS"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Context Menu

The table includes a built-in right-click context menu:

- Add row above
- Add row below
- Remove row

### Tracking Row Changes

```{r context-menu}
server <- function(input, output, session) {
  row_count <- reactiveVal(5)  # Initial rows

  observeEvent(input$my_table_edited, {
    data <- jsonlite::fromJSON(input$my_table_edited)
    new_count <- nrow(data)

    if (new_count > row_count()) {
      showNotification("Row added")
    } else if (new_count < row_count()) {
      showNotification("Row removed")
    }

    row_count(new_count)
  })
}
```

## Data Validation

### Custom Validation

```{r validation}
server <- function(input, output, session) {
  observeEvent(input$my_table_edited, {
    data <- jsonlite::fromJSON(input$my_table_edited)

    errors <- validate_data(data)

    if (length(errors) > 0) {
      showNotification(
        paste("Errors:", paste(errors, collapse = ", ")),
        type = "warning"
      )
    }
  })

  validate_data <- function(data) {
    errors <- character()

    if (any(is.na(data$name) | data$name == "")) {
      errors <- c(errors, "Name required")
    }

    if (any(data$quantity < 0, na.rm = TRUE)) {
      errors <- c(errors, "Quantity must be >= 0")
    }

    return(errors)
  }
}
```

## Export Data

### CSV Download

```{r csv}
ui <- fluidPage(
  esq_tableInput("data_table", ...),
  downloadButton("download", "Download CSV")
)

server <- function(input, output, session) {
  table_data <- reactiveVal(initial_data)

  observeEvent(input$data_table_edited, {
    table_data(jsonlite::fromJSON(input$data_table_edited))
  })

  output$download <- downloadHandler(
    filename = function() {
      paste0("data_", Sys.Date(), ".csv")
    },
    content = function(file) {
      write.csv(table_data(), file, row.names = FALSE)
    }
  )
}
```

### JSON Download

```{r json}
output$download_json <- downloadHandler(
  filename = function() {
    paste0("data_", Sys.Date(), ".json")
  },
  content = function(file) {
    jsonlite::write_json(table_data(), file, pretty = TRUE)
  }
)
```

## Auto-Save Pattern

```{r autosave}
server <- function(input, output, session) {
  auto_save <- debounce(reactive(input$my_table_edited), 2000)

  observe({
    data_json <- auto_save()
    req(data_json)

    data <- jsonlite::fromJSON(data_json)
    saveRDS(data, "auto_save.rds")

    showNotification("Auto-saved", duration = 1)
  })
}
```

## Integration with plotly

```{r plotly}
library(plotly)

ui <- fluidPage(
  fluidRow(
    column(6, esq_tableInput("data", ...)),
    column(6, plotlyOutput("chart"))
  )
)

server <- function(input, output, session) {
  table_data <- reactive({
    if (!is.null(input$data_edited)) {
      jsonlite::fromJSON(input$data_edited)
    } else {
      initial_data
    }
  })

  output$chart <- renderPlotly({
    plot_ly(table_data(), x = ~category, y = ~value, type = "bar")
  })
}
```

## Custom CSS Styling

```{r css}
ui <- fluidPage(
  tags$head(
    tags$style(HTML("
      /* Header styling */
      .esq-handsontable th {
        background: #f5f5f5;
        font-weight: 600;
      }

      /* Read-only cells */
      .esq-handsontable td.htDimmed {
        background: #fafafa;
        color: #999;
      }

      /* Invalid cells */
      .esq-handsontable td.htInvalid {
        background: #ffebee !important;
      }
    "))
  ),
  esq_tableInput("styled_table", ...)
)
```

## Keyboard Shortcuts

Built-in Handsontable shortcuts:

| Shortcut | Action |
|----------|--------|
| Arrow keys | Navigate cells |
| Enter | Edit / Move down |
| Tab | Move right |
| Shift+Tab | Move left |
| Ctrl+C | Copy |
| Ctrl+V | Paste |
| Ctrl+Z | Undo |
| Ctrl+Y | Redo |
| Delete | Clear cell |
| Escape | Cancel editing |

## Error Handling

```{r errors}
observeEvent(input$my_table_edited, {
  data <- tryCatch({
    jsonlite::fromJSON(input$my_table_edited)
  }, error = function(e) {
    showNotification("Error parsing data", type = "error")
    return(NULL)
  })

  if (is.null(data)) return()

  # Process data...
})
```

## Troubleshooting

**Table not rendering?**
- Check package installation
- Verify data frame structure

**Dropdowns empty?**
- Ensure `source` is a character vector
- Check for typos in column names

**Updates not working?**
- Verify `inputId` matches
- Check session is passed correctly

**Performance issues?**
- Reduce row count (use pagination)
- Use simpler column types
- Minimize columns
