---
title: "Column Types and Configuration"
author: "esqLABS"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Column Types and Configuration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Column Configuration Options

Each column is configured as a list with various options:

| Option | Type | Description | Default |
|--------|------|-------------|---------|
| `name` | string | Column name (required) | - |
| `type` | string | Column type (required) | - |
| `source` | vector | Options for dropdown/multiselect | NULL |
| `sortable` | boolean | Enable drag-sort for multiselect | FALSE |
| `validate` | boolean | Validate dropdown values | TRUE |
| `readOnly` | boolean | Make column read-only | FALSE |
| `width` | number | Column width in pixels | auto |

## Text Columns

Standard text input fields:

```{r text}
list(name = "description", type = "text", width = 200)

# Read-only text
list(name = "id", type = "text", readOnly = TRUE, width = 80)
```

## Numeric Columns

Number input with automatic validation:

```{r numeric}
list(name = "quantity", type = "numeric", width = 100)
list(name = "price", type = "numeric", width = 80)
```

## Checkbox Columns

Boolean checkbox cells:

```{r checkbox}
list(name = "active", type = "checkbox", width = 70)
```

Data should contain `TRUE`/`FALSE` values.

## Dropdown Columns

Single-select dropdown menus:

```{r dropdown}
list(
  name = "category",
  type = "dropdown",
  source = c("Electronics", "Clothing", "Food", "Other"),
  width = 120
)

# Allow empty values
list(
  name = "optional",
  type = "dropdown",
  source = c("", "Option A", "Option B"),
  width = 100
)

# Disable validation
list(
  name = "flexible",
  type = "dropdown",
  source = c("A", "B", "C"),
  validate = FALSE
)
```

## Multi-Select Columns

Select multiple values from a list:

```{r multiselect}
# Basic multi-select
list(
  name = "tags",
  type = "multiselect",
  source = c("urgent", "review", "approved", "pending"),
  width = 180
)

# Sortable multi-select (drag to reorder)
list(
  name = "priorities",
  type = "multiselect",
  source = c("Critical", "High", "Medium", "Low"),
  sortable = TRUE,
  width = 180
)
```

Multi-select values are stored as comma-separated strings: `"tag1, tag2, tag3"`

## Column Descriptions (Tooltips)

Add helpful tooltips to column headers:

```{r tooltips}
esq_tableInput("my_table",
  data = my_data,
  columns = my_columns,
  column_descriptions = list(
    id = "Unique identifier (auto-generated)",
    name = "Full name of the item",
    status = "Current processing status",
    priority = "Task priority level (drag to reorder)"
  )
)
```

## Complete Example

```{r complete}
library(shiny)
library(esq.handsontable)

ui <- fluidPage(
  esq_tableInput("inventory",
    data = data.frame(
      sku = c("SKU-001", "SKU-002"),
      name = c("Widget", "Gadget"),
      category = c("Electronics", "Tools"),
      price = c(29.99, 49.99),
      tags = c("sale, featured", "new"),
      in_stock = c(TRUE, FALSE),
      stringsAsFactors = FALSE
    ),
    columns = list(
      list(name = "sku", type = "text", width = 90, readOnly = TRUE),
      list(name = "name", type = "text", width = 150),
      list(name = "category", type = "dropdown",
           source = c("Electronics", "Tools", "Clothing"), width = 110),
      list(name = "price", type = "numeric", width = 80),
      list(name = "tags", type = "multiselect",
           source = c("sale", "featured", "new", "clearance"),
           sortable = TRUE, width = 150),
      list(name = "in_stock", type = "checkbox", width = 70)
    ),
    column_descriptions = list(
      sku = "Stock Keeping Unit (read-only)",
      name = "Product name",
      category = "Product category",
      price = "Price in USD",
      tags = "Product tags (drag to prioritize)",
      in_stock = "Currently in stock?"
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$inventory_edited, {
    data <- jsonlite::fromJSON(input$inventory_edited)
    print(data)
  })
}

shinyApp(ui, server)
```

## Recommended Column Widths

| Column Type | Recommended Width |
|-------------|-------------------|
| ID/Code | 60-90px |
| Short text | 100-150px |
| Long text | 200-300px |
| Numeric | 70-100px |
| Checkbox | 60-80px |
| Dropdown | 100-150px |
| Multi-select | 150-200px |
