---
title: "Conditional Cell Properties"
author: "esqLABS"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Conditional Cell Properties}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load}
library(shiny)
library(esq.handsontable)
```

## Introduction

The `cell_conditions` parameter allows you to dynamically change cell
properties based on other cell values. This enables context-aware forms
where fields become relevant or irrelevant based on user selections.

## Basic Syntax

```{r syntax}
example_condition <- list(
  list(
    column = "target_column",       # Column to modify
    when_column = "trigger_column", # Column to watch
    when_value = "trigger_value",   # Value that triggers condition
    readOnly = TRUE                 # Property to set
  )
)

str(example_condition)
```

## Simple Example

Disable fields based on a data type selection:

```{r simple}
analysis_data <- data.frame(
  name = c("Analysis 1", "Analysis 2"),
  dataType = c("Simulated", "Observed"),
  scenario = c("Baseline", ""),
  dataset = c("", "Clinical Data"),
  stringsAsFactors = FALSE
)

analysis_columns <- list(
  list(name = "name", type = "text"),
  list(name = "dataType", type = "dropdown",
       source = c("Simulated", "Observed")),
  list(name = "scenario", type = "dropdown",
       source = c("Baseline", "Treatment A", "Treatment B")),
  list(name = "dataset", type = "dropdown",
       source = c("Clinical Data", "Lab Data"))
)

analysis_conditions <- list(
  # Disable scenario when Observed
  list(
    column = "scenario",
    when_column = "dataType",
    when_value = "Observed",
    readOnly = TRUE
  ),
  # Disable dataset when Simulated
  list(
    column = "dataset",
    when_column = "dataType",
    when_value = "Simulated",
    readOnly = TRUE
  )
)

analysis_ui <- esq_tableInput(
  "analysis",
  data = analysis_data,
  columns = analysis_columns,
  cell_conditions = analysis_conditions
)

inherits(analysis_ui, "shiny.tag")
```

## Multiple Conditions on Same Column

```{r multiple}
payment_conditions <- list(
  # Disable amount for Free
  list(
    column = "amount",
    when_column = "payment",
    when_value = "Free",
    readOnly = TRUE
  ),
  # Also disable amount for Trial
  list(
    column = "amount",
    when_column = "payment",
    when_value = "Trial",
    readOnly = TRUE
  )
)

length(payment_conditions)
```

## Cascading Conditions

Disable multiple columns based on one trigger:

```{r cascade}
shipping_conditions <- list(
  # If shipping is "Pickup", disable these fields:
  list(
    column = "address",
    when_column = "shipping",
    when_value = "Pickup",
    readOnly = TRUE
  ),
  list(
    column = "delivery_date",
    when_column = "shipping",
    when_value = "Pickup",
    readOnly = TRUE
  ),
  list(
    column = "shipping_cost",
    when_column = "shipping",
    when_value = "Pickup",
    readOnly = TRUE
  )
)

vapply(shipping_conditions, `[[`, character(1), "column")
```

## Real-World Example: Clinical Trial

```{r clinical}
trial_data <- data.frame(
  arm = c("Control", "Treatment A"),
  arm_type = c("Placebo", "Active"),
  drug = c("", "Drug X"),
  dose = c(0, 100),
  frequency = c("", "Daily"),
  stringsAsFactors = FALSE
)

trial_columns <- list(
  list(name = "arm", type = "text", width = 120),
  list(name = "arm_type", type = "dropdown",
       source = c("Placebo", "Active"), width = 100),
  list(name = "drug", type = "text", width = 120),
  list(name = "dose", type = "numeric", width = 80),
  list(name = "frequency", type = "dropdown",
       source = c("Daily", "Twice Daily", "Weekly"), width = 100)
)

trial_conditions <- list(
  # Placebo arms don't need drug info
  list(column = "drug", when_column = "arm_type",
       when_value = "Placebo", readOnly = TRUE),
  list(column = "dose", when_column = "arm_type",
       when_value = "Placebo", readOnly = TRUE),
  list(column = "frequency", when_column = "arm_type",
       when_value = "Placebo", readOnly = TRUE)
)

trial_ui <- esq_tableInput(
  "trial_config",
  data = trial_data,
  columns = trial_columns,
  cell_conditions = trial_conditions
)

inherits(trial_ui, "shiny.tag")
```

## How It Works

1. When a cell value changes, all conditions are evaluated
2. If `when_column` contains `when_value`, the target `column` is modified
3. Currently, only `readOnly = TRUE` is supported
4. Disabled cells appear grayed out and cannot be edited
5. Conditions are evaluated per-row (same row only)

## Best Practices

1. **Keep conditions simple** - One value per condition
2. **Use clear trigger values** - Distinct, unambiguous options
3. **Consider defaults** - What should the initial state be?
4. **Document your logic** - Add comments explaining rules

```{r documented}
documented_conditions <- list(
  # Business rule: Observed data uses datasets, not scenarios
  list(
    column = "scenario",
    when_column = "dataType",
    when_value = "Observed",
    readOnly = TRUE
  )
)

documented_conditions[[1]]$column
```

## Limitations

- Single value matching only (no ranges or patterns)
- Only `readOnly = TRUE` property supported
- Same-row evaluation only
- Conditions evaluated on each change
