---
title: "Conditional Cell Properties"
author: "esqLABS"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Conditional Cell Properties}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

The `cell_conditions` parameter allows you to dynamically change cell properties based on other cell values. This enables context-aware forms where fields become relevant or irrelevant based on user selections.

## Basic Syntax

```{r syntax}
cell_conditions = list(
  list(
    column = "target_column",       # Column to modify
    when_column = "trigger_column", # Column to watch
    when_value = "trigger_value",   # Value that triggers condition
    readOnly = TRUE                 # Property to set
  )
)
```

## Simple Example

Disable fields based on a data type selection:

```{r simple}
esq_tableInput("analysis",
  data = data.frame(
    name = c("Analysis 1", "Analysis 2"),
    dataType = c("Simulated", "Observed"),
    scenario = c("Baseline", ""),
    dataset = c("", "Clinical Data")
  ),
  columns = list(
    list(name = "name", type = "text"),
    list(name = "dataType", type = "dropdown",
         source = c("Simulated", "Observed")),
    list(name = "scenario", type = "dropdown",
         source = c("Baseline", "Treatment A", "Treatment B")),
    list(name = "dataset", type = "dropdown",
         source = c("Clinical Data", "Lab Data"))
  ),
  cell_conditions = list(
    # Disable scenario when Observed
    list(
      column = "scenario",
      when_column = "dataType",
      when_value = "Observed",
      readOnly = TRUE
    ),
    # Disable dataset when Simulated
    list(
      column = "dataset",
      when_column = "dataType",
      when_value = "Simulated",
      readOnly = TRUE
    )
  )
)
```

## Multiple Conditions on Same Column

```{r multiple}
cell_conditions = list(
  # Disable amount for Free
  list(
    column = "amount",
    when_column = "payment",
    when_value = "Free",
    readOnly = TRUE
  ),
  # Also disable amount for Trial
  list(
    column = "amount",
    when_column = "payment",
    when_value = "Trial",
    readOnly = TRUE
  )
)
```

## Cascading Conditions

Disable multiple columns based on one trigger:

```{r cascade}
cell_conditions = list(
  # If shipping is "Pickup", disable these fields:
  list(
    column = "address",
    when_column = "shipping",
    when_value = "Pickup",
    readOnly = TRUE
  ),
  list(
    column = "delivery_date",
    when_column = "shipping",
    when_value = "Pickup",
    readOnly = TRUE
  ),
  list(
    column = "shipping_cost",
    when_column = "shipping",
    when_value = "Pickup",
    readOnly = TRUE
  )
)
```

## Real-World Example: Clinical Trial

```{r clinical}
esq_tableInput("trial_config",
  data = data.frame(
    arm = c("Control", "Treatment A"),
    arm_type = c("Placebo", "Active"),
    drug = c("", "Drug X"),
    dose = c(0, 100),
    frequency = c("", "Daily")
  ),
  columns = list(
    list(name = "arm", type = "text", width = 120),
    list(name = "arm_type", type = "dropdown",
         source = c("Placebo", "Active"), width = 100),
    list(name = "drug", type = "text", width = 120),
    list(name = "dose", type = "numeric", width = 80),
    list(name = "frequency", type = "dropdown",
         source = c("Daily", "Twice Daily", "Weekly"), width = 100)
  ),
  cell_conditions = list(
    # Placebo arms don't need drug info
    list(column = "drug", when_column = "arm_type",
         when_value = "Placebo", readOnly = TRUE),
    list(column = "dose", when_column = "arm_type",
         when_value = "Placebo", readOnly = TRUE),
    list(column = "frequency", when_column = "arm_type",
         when_value = "Placebo", readOnly = TRUE)
  )
)
```

## How It Works

1. When a cell value changes, all conditions are evaluated
2. If `when_column` contains `when_value`, the target `column` is modified
3. Currently, only `readOnly = TRUE` is supported
4. Disabled cells appear grayed out and cannot be edited
5. Conditions are evaluated per-row (same row only)

## Best Practices

1. **Keep conditions simple** - One value per condition
2. **Use clear trigger values** - Distinct, unambiguous options
3. **Consider defaults** - What should the initial state be?
4. **Document your logic** - Add comments explaining rules

```{r documented}
cell_conditions = list(
  # Business rule: Observed data uses datasets, not scenarios
  list(
    column = "scenario",
    when_column = "dataType",
    when_value = "Observed",
    readOnly = TRUE
  )
)
```

## Limitations

- Single value matching only (no ranges or patterns)
- Only `readOnly = TRUE` property supported
- Same-row evaluation only
- Conditions evaluated on each change
