---
title: "Dynamic Updates"
author: "esqLABS"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dynamic Updates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

The `updateEsqTable()` function allows you to update dropdown options at runtime without recreating the table.

## Syntax

```{r syntax}
updateEsqTable(
  session,           # Shiny session object
  inputId,           # Table input ID
  options = list()   # Named list of column options
)
```

## Basic Example

```{r basic}
library(shiny)
library(esq.handsontable)

ui <- fluidPage(
  selectInput("region", "Region:",
    choices = c("North America", "Europe", "Asia")),
  actionButton("update", "Update Countries"),

  esq_tableInput("sales",
    data = data.frame(
      product = c("Widget", "Gadget"),
      country = c("USA", "Canada")
    ),
    columns = list(
      list(name = "product", type = "text"),
      list(name = "country", type = "dropdown",
           source = c("USA", "Canada", "Mexico"))
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$update, {
    countries <- switch(input$region,
      "North America" = c("USA", "Canada", "Mexico"),
      "Europe" = c("UK", "Germany", "France", "Spain"),
      "Asia" = c("Japan", "China", "South Korea", "India")
    )

    updateEsqTable(session, "sales",
      options = list(country = countries)
    )
  })
}

shinyApp(ui, server)
```

## Update Multiple Columns

```{r multiple}
updateEsqTable(session, "my_table",
  options = list(
    country = c("USA", "Canada"),
    state = c("California", "Texas"),
    city = c("Los Angeles", "Houston")
  )
)
```

## Reactive Updates

```{r reactive}
server <- function(input, output, session) {
  available_options <- reactive({
    # Could be a database query
    switch(input$category,
      "Electronics" = c("Phones", "Laptops", "Tablets"),
      "Clothing" = c("Shirts", "Pants", "Shoes"),
      c("Other")
    )
  })

  observe({
    updateEsqTable(session, "product_table",
      options = list(subcategory = available_options())
    )
  })
}
```

## Cascading Dropdowns

```{r cascading}
location_data <- list(
  "USA" = c("California", "Texas", "New York"),
  "Canada" = c("Ontario", "Quebec", "BC")
)

server <- function(input, output, session) {
  observeEvent(input$country, {
    states <- location_data[[input$country]]

    updateEsqTable(session, "location_table",
      options = list(state = states)
    )
  })
}
```

## Database-Driven Updates

```{r database}
server <- function(input, output, session) {
  observeEvent(input$refresh, {
    # Query from database
    # categories <- dbGetQuery(con,
    #   "SELECT DISTINCT category FROM products")

    categories <- c("New Category 1", "New Category 2")

    updateEsqTable(session, "product_table",
      options = list(category = categories)
    )

    showNotification("Options updated!", type = "message")
  })
}
```

## Shiny Modules

When using modules, the session is automatically namespaced:

```{r modules}
tableModuleServer <- function(id) {
  moduleServer(id, function(input, output, session) {
    observeEvent(input$refresh, {
      # Use inputId without namespace prefix
      updateEsqTable(session, "data_table",
        options = list(category = new_categories)
      )
    })
  })
}
```

## Performance Tips

### Batch Updates

```{r batch}
# Good: Single call
updateEsqTable(session, "table",
  options = list(
    col1 = opts1,
    col2 = opts2,
    col3 = opts3
  )
)

# Avoid: Multiple calls
# updateEsqTable(session, "table", options = list(col1 = opts1))
# updateEsqTable(session, "table", options = list(col2 = opts2))
```

### Debouncing

```{r debounce}
server <- function(input, output, session) {
  search_debounced <- debounce(reactive(input$search), 300)

  observe({
    term <- search_debounced()
    options <- fetch_options(term)
    updateEsqTable(session, "table",
      options = list(items = options)
    )
  })
}
```

## Best Practices

1. **Update only when needed** - Don't update on every keystroke
2. **Batch updates** - Combine multiple columns in one call
3. **Provide feedback** - Show notifications for slow updates
4. **Handle invalid values** - Warn users about orphaned values
5. **Test edge cases** - Empty options, special characters
