---
title: "Getting Started with esq.handsontable"
author: "esqLABS"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with esq.handsontable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

`esq.handsontable` is a powerful R Shiny package that provides an interactive, Excel-like data table component. Built on [Handsontable](https://handsontable.com/) and React, it offers features like dropdowns, multi-select, validation, conditional cell disabling, and more.

## Installation

```{r install}
# Install from GitHub
devtools::install_github("esqlabs/esq.handsontable")
```

## Quick Start

Here's a minimal example:

```{r quick-start}
library(shiny)
library(esq.handsontable)

ui <- fluidPage(
  esq_tableInput("my_table",
    data = data.frame(
      name = c("Alice", "Bob", "Charlie"),
      age = c(25, 30, 35),
      active = c(TRUE, FALSE, TRUE)
    ),
    columns = list(
      list(name = "name", type = "text"),
      list(name = "age", type = "numeric"),
      list(name = "active", type = "checkbox")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$my_table_edited, {
    data <- jsonlite::fromJSON(input$my_table_edited)
    print(data)
  })
}

shinyApp(ui, server)
```

## Core Concepts

### The `esq_tableInput()` Function

| Argument | Description |
|----------|-------------|
| `inputId` | Unique identifier for the table |
| `data` | A data frame with initial data |
| `columns` | List of column configurations |
| `cell_conditions` | Optional conditional cell rules |
| `column_descriptions` | Optional tooltips for columns |

### Column Types

| Type | Description |
|------|-------------|
| `text` | Standard text input |
| `numeric` | Number input with validation |
| `checkbox` | Boolean checkbox |
| `dropdown` | Single-select dropdown |
| `multiselect` | Multi-select dropdown |

### Retrieving Data

```{r retrieve}
observeEvent(input$my_table_edited, {
  data <- jsonlite::fromJSON(input$my_table_edited)
  # Use the data...
})
```

## Next Steps

- See `vignette("column-types")` for detailed column configuration
- See `vignette("conditional-cells")` for dynamic cell behavior
- See `vignette("dynamic-updates")` for updating options at runtime
- Check the `examples/` folder for complete working apps
